### 2025-09-25 mixed-models.qmd
20250925 - cross-classification

Another form of nonindependent data is when data are *cross-classified* (also called crossed).
Data are cross-classified when the lower-level units are classified by two or more upper-level units, but the upper-level units are not hierarchical or nested within one another.
For instance, children may be nested within the crossing of schools and neighborhoods.
That is, children are nested within schools, and children are also nested within neighborhoods; however, children attending the same school may not necessarily be from the same neighborhood, and children from the same neighborhood may not necessarily attend the same school.
The data would still have a two-level hierarchy, but the data would be nested in specific school-neighborhood combinations.
That is, children are nested within the cross-classification of schools and neighborhoods.

Applied to longitudinal fantasy football data, Most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players, with the player's ID as the cluster variable.
However, you could get more complicated.
For instance, you could estimate a three-level model of timepoints nested within players nested within teams.
This would help account for players' productivity benefitting or taking a hit from being on a given team, and it would account for players changing teams.
Alternatively, consistent with @Usami2018, you could estimate a cross-classified model of players' individual observations/measurement occasions belonging to two upper-level units: players and time points (e.g., week 1, week 2, 2025 season, etc.).
This would help account for systematic effects that are specific to a given timepoint (e.g., week or season).

>>>>

Another form of nonindependent data is when data are *cross-classified* (also called crossed).
Data are cross-classified when the lower-level units are classified by two or more upper-level units, but the upper-level units are not hierarchical or nested within one another.
For instance, children may be nested within the crossing of schools and neighborhoods.
That is, children are nested within schools, and children are also nested within neighborhoods; however, children attending the same school may not necessarily be from the same neighborhood, and children from the same neighborhood may not necessarily attend the same school.
The data would still have a two-level hierarchy, but the data would be nested in specific school-neighborhood combinations.
That is, children are nested within the cross-classification of schools and neighborhoods.
An example of cross-classified data in football is that players are nested within teams and within agents.
Agents are not nested within teams, and teams are not nested within agents.
That is, teams are nested within the cross-classification of teams and agents.

Applied to longitudinal fantasy football data, most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players, with the player's ID as the cluster variable.
However, you could get more complicated.
For instance, you could estimate a three-level model of timepoints nested within players nested within teams.
This would help account for players' productivity benefitting or taking a hit from being on a given team, and it would account for players changing teams.
Alternatively, consistent with @Usami2018, you could estimate a cross-classified model of players' individual observations/measurement occasions belonging to two upper-level units: players and time points (e.g., week 1, week 2, 2025 season, etc.).
This would help account for systematic effects that are specific to a given timepoint (e.g., week or season).
